Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-07-30 Manuscript/MPESubmission/Revision/response.tex

% TG: I'm confused for the second part of the question:
% TG: In figure 4 and in Table 1, if you look at the Bayesian consensus (grey) trees and the Bayesian posterior trees (orange) for RF, they do not really overlap. Ever. The 0.11 probability must come from the parameters combinations that, when is really shitty must show a little bit of overlap. But for each parameter one by one: nope. No overlap.
% TG: On the other side, in figure 4 and Table 1, for Bayesian consensus and Bayesian posterior for Tr, the do overlap much more. And that's what's in the Table 1 (min 0.21; median 0.56; and max 0.87).
% TG: Did the reviewer just miss read the labels of the Table 1 or am I miles away from what he means?

\item{\textcolor{blue}{\textbf{Results - Bayesian Triplets:} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
Values in Figures 4 and 5 clearly drop below zero, although negative values are not illustrated on the axis, and this indicates that the tree topologies are no better than, or even worse than, random trees.
It is especially true when fossils are missing data, and that is going to be an important finding with practical application but must be addressed. }}

We modified the figures 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
% TG: a a paragraph in the results about the wild-card fossils
As mentioned in the previous comment (reviewer 1 specific comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
This has two related effects.
The first one being to retain "good" clade conservation (i.e. avoiding "false-positive" nodes by collapsing unsupported ones).
And the second one is that when these clades are collapsed, it reduces the amount of available triplets and thus increases the "wild-card taxa" score (Normalised Triplets metric).
In fact, when a clade ((a;b);c); is collapsed to (a;b;c), the whole clade is conserved (a, b and c are still in a common clade) however they also all become wild-card taxa: the position of a regarding b and c is now incorrect.

@@

\item{\textcolor{blue}{\textbf{Figure 4:} Figure 4 has a color scheme issue, the colors in the caption do not match up to the figure.}}

We apologies for this silly mistake and fixed the colors in the figure. We also changed the color scheme to match with the general color scheme of the paper (black, grey, orange and blue).

\item{\textcolor{blue}{\textbf{Figure 4-5:} Am I wrong or shouldn't the y-axis for Normalised Triplets metric go below 0 - it looks those metrics for many missing data combinations have values below zero indicating trees are more different from the 'best' tree than expected by chance (Appendix B 1.1., see Appendix C Tables 1-3) and that is important to show.}}

We fixed this graphical mistake as state above in reviewer 1 specific comment 13.

@@

\item{\textcolor{blue}{\textbf{line 515:} '…the size of the matrix' missing 'of'}}

We fixed this typo.

\item{\textcolor{blue}{\textbf{line 518 to 520:} I am not sure that (Wagner 2000) states that incongruence between molecular and morphological data is 'more important in small morphological matrices'. }}

We now refer to two more classical citations (Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al. 1993 (Annual Review of Ecology and Systematics)) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to small sized matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters).

@@

\item{\textcolor{blue}{\textbf{line 528:} missing close parenthesis after Ni et al.}}

We fixed this typo.

\item{\textcolor{blue}{\textbf{line 532:} Wright and Hillis 2014 found that increasing the size of the dataset improves topological accuracy, and they do not argue that simply increasing the number of characters increases the amount of homoplasy. Rather they found that the fastest evolving characters exhibited homoplasy, while the slowest evolving lacked phylogenetic signal. The citation seems misrepresented. Further, there was higher phylogenetic accuracy of simulated datasets with more characters than fewer characters, for slow or fast evolving characters (Wiens 2005).}}

We retracted the entire statement concerning homoplasy and matrix size ("Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters (Wright and Hillis 2014).")

@@

\item{\textcolor{blue}{\textbf{line 540:} replace 'a lot' with 'larger proportions of missing data', a lot seems too colloquial for this kind of article. }}

We modified the sentence as suggested: "they have a larger proportion of missing data" (line @@@).

\item{\textcolor{blue}{\textbf{lines 542-547:} These seem like the most important practical results and should be more strongly emphasized. These are very important results for practitioners and the authors should try to make it more clear to the reader that this is a key take-home message.}}

We modified the paragraph to:

"Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combinations of the other parameters ($M_F$ and $N_C$; Figure 6).
In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) doesn't matter and and increase/decrease in both does not affect topology.
A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below)." lines (@@@)

@@

\item{\textcolor{blue}{\textbf{line 557-558:} '…than the "best" maximum likelihood tree'}}

We modified the sentence as suggested : "than the ``best'' Maximum Likelihood tree" (line @@@).

\item{\textcolor{blue}{\textbf{line 549-565:} again, I am left wondering how much the differences in performance of max. likelihood and Bayesian inference techniques are driven by the use of the 'true' tree as a starting tree in the Bayesian MCMC search, while the likelihood analysis had no such guide tree to begin with. The authors at least need to make this point clear, if not re-run some subset of analyses without introducing a starting tree to test the effects of using the 'true' tree as a starting tree or not. I understand the computational limitations, but the practical limitation that empiricists will never have the 'true' tree to start with may (or may not) have a large impact on our interpretation and the generality of these findings. }}

As developed above (reviewer 1, specific comment 29), we do understand the reviewers concern and added an analysis in the Appendix A (figure A.2 and tables A.1 and A.2) where we tested the effect of using the ``true'' tree as a starting tree. We found that there is no significant effect between using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.

We refer to this supplementary analysis in the main text as follows:

"Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference)." lines @@@

@@

\item{\textcolor{blue}{\textbf{line 612-613:} too obvious and not especially helpful. We should aim to have as little missing data as possible…more helpful, and something especially discovered in this study, is concentrating on having overlapping morphological characters between the living and fossil taxa to be able to place the living taxa with strong support via large datasets (combined data) and being able to position the fossils based on their shared derived characters with living taxa. This is an important point to make clearer. }}

We are thankful for this comment as, indeed, our suggestion was a bit to obvious and not really helpful. We modified the end of this paragraph as follows:

"Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e. 50\%) for accurately combining both living and fossil species in phylogenies.
Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data."

%Original: Therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.

@@

\item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}

We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 specific comment 50)
%We modified the sentence as follow: "Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa." (line @@@), TG: actually we removed the sentence.

@@

Unfortunately, we didn't found many studies formally testing the important questions asked by the reviewer. However, there exist few empirical analysis that contain results supporting our claims.
Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree. Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95). We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rates estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.

We modified the mentioned section of the conclusion as follows:
"Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
This has also been observed in empirical data (e.g. Arcila et al., 2015) were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds (e.g. Slater, 2013), taxonomic constraints (e.g. Slater, 2013; Beck and Lee, 2014) or even by fixing the topology (e.g. Ronquist et al., 2012a).
Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes."
%Original: Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the “best” tree topology than the Maximum Likelihood trees. Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses. The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data. Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.

@@

\item{\textcolor{blue}{\textbf{line 628:} use of computer cluster may be better placed in the methods, depending on the norm for MPE.}}

The Lonsdale cluster requires this specific line in the acknowledgements of any study using it's infrastructure (\url{http://www.tchpc.tcd.ie/resources/acknowledgementpolicy}).
If requested, however, we can also give additional details on the cluster used in the methods section.
The current version of the manuscript only contains information on the nodes clock speed (2.30GHz clock speed nodes) lines @@@.

@@

We apologies for this silly mistake and thanks the reviewer for pointing it out. We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".

\item{\textcolor{blue}{\textbf{Figure 5:} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic.
But concerning the Normalized Triplets metric, according to the text and appendix, if this index is 0, the similarity of the missing data tree to the best tree is no more than random, and < 0 it is more different than random.
For the Bayesian analyses, many results had values < 0 but the axis is set to 0.
Why not illustrate the distribution of values that is < 0? %Fixed
From this result, it appears the Bayesian analysis does quite poorly with missing data in terms of rouge taxa, and worse than max. likelihood.
This figure also suggests to me that the methods perform pretty badly with missing data, in contrast to previous findings that missing data are not overly deleterious.
For example, when all data are complete except Mf=50\%, the normalized RF metric is <0.8.
Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005), although for small datasets such as the one used in this study, the results show a similar decline in accuracy.
This is a point that should be expanded upon in the discussion - how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?}}

We thank the reviewer for this excellent suggestion and modified our graphical abstract. We also fixed the Normalised Triplets metrics axis to range from -0.5 to 1. We added a horizontal doted line at Triplets = 0 to make sure the reader take notes is this change of scale.

%how these results compare to other missing data studies
%Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005)
%In Wiens when missing data ~50\%, accuracy is ~70/60

%what do they tell us about the performance of the inference methods with missing data?
%Bayesian seems more conservative than Maximum Likelihood.
%Better to be more conservative?

% STILL TO DO

% TG: maybe put that somewhere?
% In this study we chose to differentiate both metrics (RF and Triplets) because they account for two important aspects of topology in macroevolution.
% The RF metric is a proxy for clade conservation which is important for macroevolutionary or macroecological methods such as phylogenetic comparative methods (rather than the specific position of each taxa, the conservation and support of clades is essential).
% For example, for a tree (((a;b);c)((d;e);f);, swapping the taxa a and c (the clade is conserved) as less impact on comparative methods than swapping a and d (the clades are less conserved). %TG: is this kind of right? For PGLS for example...
% % Or, other example with the same tree but for palaebioography: if ((a;b);c); and ((d;e);f); have different geographical origin; then a swap a and d will have important implications.
% On the other hand the Triplets metric might be more important from macroevolutionary methods such as dating.
% For example on the same tree (((a;b);c)((d;e);f); if a is a younger taxa than c, the swap of a and c will have a bigger effect on the age of the clade.

### 2015-07-30 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated revisions

Additionally, the number of morphological characters determines the size the of matrix.
This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
The sizes of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}. 
Nonetheless, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{kv} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

\subsection{Combined effect of missing data parameters}
As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when they have a larger proportion of missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Figure 5).
It is important, however, to notice that the effect of each parameter is not additive.
Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
It seems therefore more important to integrate a maximum of living taxa with morphological data as well as a maximum number of characters.
Additionally, when both $M_{L}$ reach 50\% missing data and $N_{C}$ reach 50 characters, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).

\subsection{Effects of tree inference methods}
Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A - @@@ for details on the analysis of the effect of the starting tree in Bayesian inference).
As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
This may also be because the ``best'' Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best'' Maximum Likelihood tree.
Nonetheless, we minimized the probability of unresolved ``best'' trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).

@@

Another practical implication of our results regards the tree inference methods.
Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating \citep[e.g.][; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution]{ronquista2012,Wood01032013,BEASTmaster}.
Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e. creating a ``false positive'' node).
Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g. \citealt{FritzTree} and \citealt{jetzthe2012} trees used in \citealt{healy2014}).

\section{Conclusions}
Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
Therefore we advise filling as many gaps in Total Evidence matrices as possible.
Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa.

Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
This has also been observed in empirical data \cite[e.g.][]{Arcila2015131} were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds \citep[e.g.][]{slaterphylogenetic2013}, taxonomic constraints \citep[e.g.][]{slaterphylogenetic2013,beckancient2014} or even by fixing the topology \cite[e.g.][]{ronquista2012}.
Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.

The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.

% Do NOT remove this, even if you are not including acknowledgments.
\section{Acknowledgments}
Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript. Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster. Thanks to the two anonymous reviewers for their useful and enthusiastic %or "encouraging"
comments. Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland. This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

@@

\bigskip
\noindent
\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange). The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.

>>>>

Additionally, the number of morphological characters determines the size the of matrix.
This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
\ul{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
Therefore, our simulations reflect realistic parameters.
Nonetheless, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{kv} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

\subsection{Combined effect of missing data parameters}
As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when \hl{they have a larger proportion of missing data} (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Figure 5).
It is important, however, to notice that the effect of each parameter is not additive.
\hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combinations of the other parameters ($M_F$ and $N_C$; Figure 6).
In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) doesn't matter and and increase/decrease in both does not affect topology.
A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below).}

\subsection{Effects of tree inference methods}
Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over \hl{parsimony} methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
\hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference).}
As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
This may also be because the ``best'' Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best'' Maximum Likelihood tree.
Nonetheless, we minimized the probability of unresolved ``best'' trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).

@@

\hl{Another practical implication of our results regards the tree inference methods.
Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating} \citep[\hl{e.g.}][\hl{; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution}]{ronquista2012,Wood01032013,BEASTmaster}.
\hl{Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e. creating a ``false positive'' node).
Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g. }\citealt{FritzTree}\hl{ and }\citealt{jetzthe2012}\hl{ trees used in }\citealt{healy2014}\hl{).}

\section{Conclusions}
Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
\hl{Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e. 50\%) for accurately combining both living and fossil species in phylogenies.
Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data.}

\hl{Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
This has also been observed in empirical data} \cite[\hl{e.g.}][]{Arcila2015131}\hl{ were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even by fixing the topology }\cite[\hl{e.g.}][]{ronquista2012}.
\hl{Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.}

\hl{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.}

\section{Acknowledgments}
Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript. Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster. Thanks to the two anonymous reviewers for their useful and enthusiastic % TG: or "useful and encouraging"? I feel just useful does gives justice to them, especially to the work of reviewer 1...
comments. Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland. This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

@@

\bigskip
\noindent
\textbf{Figure 4:} The effects of increasing missing data on topological recovery using \hl{Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)}. The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
